3D Magnetic modelling of ellipsoidal bodies

### 2017-02-17 manuscript/manuscript.tex
remove the theta and describe the magnetization in the main coordinate system

Consider a magnetized ellipsoid immersed in
a uniform magnetic field $\mathbf{H}_{0}$ (in $\unit{Am^{-1}}$)
given by
\begin{equation}
\mathbf{H}_{0} = \| \mathbf{H}_{0} \| \left[
\begin{array}{c}
\cos I \: \cos D \\
\cos I \: \sin D \\
\sin I
\end{array}
\right] \: ,
\label{eq:H0}
\end{equation}
where $\| \cdot \|$ denotes the Euclidean norm and
$D$ and $I$ are respectively, the declination and inclination of the 
local-geomagnetic field in the main coordinate system
(Fig. \ref{fig:orientation-angles}a).
This uniform field can be, for example, the main component of the
Earth's magnetic field, which is usually assumed to be generated
by the Earth's liquid core.
In the absence of conduction currents, 
the total magnetic field $\mathbf{H}(\mathbf{r})$
at the position $\mathbf{r}$ (Eqs. \ref{eq:A} and 
\ref{eq:coord_transformation}) of a point referred to the main 
coordinate system is defined as follows \citep{sharma1966, 
eskola1980, reitz1992, sttraton2007}:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0} - \nabla V(\mathbf{r}) \: ,
\label{eq:H}
\end{equation}
where the second term is the negative gradient of 
the magnetic scalar potential $V(\mathbf{r})$ given by:
\begin{equation}
V(\mathbf{r}) = -\frac{1}{4\pi} \iiint_{\vartheta} 
\mathbf{M}(\mathbf{r}^{\prime})^{\top} 
\nabla \left(
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\right) \, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:phi-potential}
\end{equation}
In this equation, $\mathbf{r}^{\prime} = [\begin{array}{ccc} 
x^{\prime} & y^{\prime} & z^{\prime} \end{array} ]^{\top}$
is the position vector of a point located within the volume $\vartheta$, 
the integral is conducted over the variables 
$x^{\prime}$, $y^{\prime}$ and, $z^{\prime}$ and 
$\mathbf{M}(\mathbf{r}^{\prime})$ is the magnetization vector
(in $\unit{Am^{-1}}$).
Eq. \ref{eq:phi-potential} is valid anywhere, 
independently if the position vector $\mathbf{r}$ represents
a point located inside or outside the magnetized body
\citep{dubois1896, reitz1992, sttraton2007}.

@@

If the principal susceptibilities are different from
each other, we say that the body has an
anisotropy of magnetic susceptibility (AMS).
The AMS is generally associated to the preferred orientation
of the grains of magnetic minerals forming the rock
\citep{fuller1963, uyeda1963, janak1972, hrouda1982, 
thompson1986, macdonald1987, rochette1992, dunlop1997,
tauxe2003}.
For the particular case in which the principal directions 
coincide with the ellipsoid axes, the matrix $\mathbf{U}$ is
equal to the matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
depending on the ellipsoid type.
Another important particular case is that in which the
susceptibility is isotropic and, consequently, the principal
susceptibilities $k_{1}$, $k_{2}$, and $k_{3}$ (Eq. \ref{eq:K})
are equal to a constant $\chi$. In this case, the susceptibility
tensor $\mathbf{K}$ (Eq. \ref{eq:K}) assumes the particular form
\begin{equation}
\mathbf{K} = \chi \, \mathbf{I} \: ,
\label{eq:K-isotropic}
\end{equation}
where $\mathbf{I}$ represents the identity matrix.

@@

Let us pre-multiply the uniform internal field $\tilde{\mathbf{H}}^{\dagger}$
(Eq. \ref{eq:Hi-tilde}) by the transformed 
susceptibility tensor $\tilde{\mathbf{K}}$ 
(Eq. \ref{eq:H-tilde}) to obtain
\begin{equation}
\begin{split}
\tilde{\mathbf{M}} 
&= \tilde{\mathbf{K}} 
\left( \mathbf{I} - \tilde{\mathbf{N}}^{\dagger} \, \tilde{\mathbf{K}} \right)^{-1}
\tilde{\mathbf{H}}_{0} \\
&=  
\left( \mathbf{I} - \tilde{\mathbf{K}} \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\tilde{\mathbf{K}} \, \tilde{\mathbf{H}}_{0}
\end{split} \: ,
\label{eq:M-tilde}
\end{equation}
where $\tilde{\mathbf{M}}$
represents the transformed magnetization, as can be easily
verified by using Eqs. \ref{eq:M-KH} and \ref{eq:H-tilde}.
The matrix identity used for obtaining the second line of Eq. 
\ref{eq:M-tilde} is given by \citet[p. ~151]{searle1982}.

Equation \ref{eq:M-tilde} can be easily generalized for the case in 
which the ellipsoid has also a uniform remanent magnetization
$\tilde{\mathbf{M}}_{R}$. Let us first consider that
the uniform remanent magnetization satisfies the condition
$\tilde{\mathbf{H}}_{A} = \tilde{\mathbf{K}}^{-1} \tilde{\mathbf{M}}_{R}$,
where $\tilde{\mathbf{H}}_{A}$ represents a hypothetical 
uniform ancient field. Then, if we assume that $\tilde{\mathbf{H}}_{0}$,
in Eqs. \ref{eq:Hi-tilde} and \ref{eq:M-tilde},
is in fact the sum of the uniform magnetic field $\tilde{\mathbf{H}}_{0}$ 
and the hypothetical ancient field $\tilde{\mathbf{H}}_{A}$, we obtain the 
following generalized equation
\begin{equation}
\tilde{\mathbf{M}} =  
\left( \mathbf{I} - \tilde{\mathbf{K}} \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\left( \tilde{\mathbf{K}} \, \tilde{\mathbf{H}}_{0} + \tilde{\mathbf{M}}_{R} \right) \: .
\label{eq:M-tilde-remanence}
\end{equation}
Equation \ref{eq:M-tilde-remanence} is consistent with that given by
\citet[Eq. ~38]{clark1986}. It shows the combined effect of the
anisotropy of magnetic susceptibility, which is represented by the 
susceptibility tensor $\tilde{\mathbf{K}}$ and reflects the
preferred orientation of the magnetic minerals forming the
body, and the shape anisotropy, which is represented by the
depolarization tensor $\tilde{\mathbf{N}}^{\dagger}$ and
reflects the self-demagnetization associated to the body shape.
Notice that the resultant magnetization $\tilde{\mathbf{M}}$
(Eq. \ref{eq:M-tilde-remanence}) does not necessarily have
the same direction as the inducing field $\tilde{\mathbf{H}}_{0}$.
The angular difference $\theta$ between the resultant magnetization
and the inducing field is the same, whether the quantities
are defined in the main coordinate system (Fig. \ref{fig:orientation-angles}a)
or local coordinate system (Fig. \ref{fig:orientation-angles}b).
For convenience, we define $\theta$ (in degrees) in the main coordinate system
by using the scalar product definition as follows:
\begin{equation}
\theta = \frac{180}{\pi} \, \cos^{-1} \left( 
\frac{\mathbf{M}^{\top}\mathbf{H}_{0}}
{\| \mathbf{M} \| \; \| \mathbf{H}_{0} \|}\right) \: ,
\label{eq:theta}
\end{equation}
where $\mathbf{M}$ is obtained by pre-multiplying the
magnetization $\tilde{\mathbf{M}}$ (Eq. \ref{eq:M-tilde-remanence})
by the orthogonal matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
which is defined according to the ellipsoid type, and
$\mathbf{H}_{0}$ is given by Eq. \ref{eq:H0}.

For the particular case in which the susceptibility is isotropic,
the susceptibility tensor is defined according to Eq. \ref{eq:K-isotropic}.
In this case, the magnetization $\mathbf{M}$ (Eq. \ref{eq:M-KH}),
referred to the main coordinate system (Fig. \ref{fig:orientation-angles}a),
can be written as follows:
\begin{equation}
\mathbf{M} = \mathbf{\Lambda} \left( \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \right) \: ,
\label{eq:M-K-isotropic}
\end{equation}
where
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left[ \begin{array}{ccc}
\frac{1}{1 - \chi \, \tilde{n}^{\dagger}_{11}} &  &  \\
 & \frac{1}{1 - \chi \, \tilde{n}^{\dagger}_{22}} &  \\
 &  & \frac{1}{1 - \chi \, \tilde{n}^{\dagger}_{33}}
\end{array} \right] \mathbf{V}^{\top} \: .
\label{eq:Lambda}
\end{equation}
This equation is in perfect agreement with those presented 
by \citet[Eqs. ~13--15]{guo2001}.
The first term, depending on the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}), 
represents the induced magnetization whereas the term depending on
$\mathbf{M}_{R}$ is the remanent magnetization.
Equation \ref{eq:M-K-isotropic} reveals that, as pointed out by
many authors
\citep[e.g.,][]{maxwell1873, dubois1896, stoner1945, clark1986, sttraton2007},
the induced magnetization opposes the inducing field 
if it is parallel to an ellipsoid axis, 
independently of the ellipsoid type. Otherwise, the
magnetization is not necessarily parallel to
the inducing field.
If we additionally consider that $\chi << 1$, the matrix 
$\mathbf{\Lambda}$ (Eqs. \ref{eq:M-K-isotropic} and \ref{eq:Lambda})
approaches to the identity and the magnetization
$\mathbf{M}$ (Eqs. \ref{eq:M-KH} and \ref{eq:M-K-isotropic})
can be approximated as follows:
\begin{equation}
\mathbf{M} \approx \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-approx}
\end{equation}
which is the classical equation describing the resultant
magnetization in applied geophysics \citep[p. ~89]{blakely1996}.
Notice that, in this particular case, the induced magnetization
is parallel to the inducing field $\mathbf{H}_{0}$ (Eq. \ref{eq:H0}),
whether it is parallel to an ellipsoid axis or not.
Usually, Eq. \ref{eq:M-approx}
is considered a good approximation for $\chi \leq 0.1$ SI.

>>>>

Consider a magnetized ellipsoid immersed in
a uniform magnetic field $\mathbf{H}_{0}$ (in $\unit{Am^{-1}}$)
given by
\begin{equation}
\mathbf{H}_{0} = \| \mathbf{H}_{0} \| \left[
\begin{array}{c}
\cos I \: \cos D \\
\cos I \: \sin D \\
\sin I
\end{array}
\right] \: ,
\label{eq:H0}
\end{equation}
where $\| \cdot \|$ denotes the Euclidean norm and
$D$ and $I$ are respectively, the declination and inclination of the 
local-geomagnetic field in the main coordinate system
(Fig. \ref{fig:orientation-angles}a).
This uniform field can be, for example, the main component of the
Earth's magnetic field, which is usually assumed to be generated
by the Earth's liquid core.
In the absence of conduction currents, 
the total magnetic field $\mathbf{H}(\mathbf{r})$
at the position $\mathbf{r}$ (Eqs. \ref{eq:A} and 
\ref{eq:coord_transformation}) of a point referred to the main 
coordinate system is defined as follows \citep{sharma1966, 
eskola1980, reitz1992, sttraton2007}:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0} - \nabla V(\mathbf{r}) \: ,
\label{eq:H}
\end{equation}
where the second term is the negative gradient of 
the magnetic scalar potential $V(\mathbf{r})$ given by:
\begin{equation}
V(\mathbf{r}) = -\frac{1}{4\pi} \iiint_{\vartheta} 
\mathbf{M}(\mathbf{r}^{\prime})^{\top} 
\nabla \left(
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\right) \, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:V-potential}
\end{equation}
In this equation, $\mathbf{r}^{\prime} = [\begin{array}{ccc} 
x^{\prime} & y^{\prime} & z^{\prime} \end{array} ]^{\top}$
is the position vector of a point located within the volume $\vartheta$, 
the integral is conducted over the variables 
$x^{\prime}$, $y^{\prime}$ and, $z^{\prime}$ and 
$\mathbf{M}(\mathbf{r}^{\prime})$ is the magnetization vector
(in $\unit{Am^{-1}}$).
Eq. \ref{eq:V-potential} is valid anywhere, 
independently if the position vector $\mathbf{r}$ represents
a point located inside or outside the magnetized body
\citep{dubois1896, reitz1992, sttraton2007}.

@@

If the principal susceptibilities are different from
each other, we say that the body has an
anisotropy of magnetic susceptibility (AMS).
The AMS is generally associated to the preferred orientation
of the grains of magnetic minerals forming the rock
\citep{fuller1963, uyeda1963, janak1972, hrouda1982, 
thompson1986, macdonald1987, rochette1992, dunlop1997,
tauxe2003}.
For the particular case in which the principal directions 
coincide with the ellipsoid axes, the matrix $\mathbf{U}$ is
equal to the matrix $\mathbf{V}$ (Eq. \ref{eq:A}).
Another important particular case is that in which the
susceptibility is isotropic and, consequently, the principal
susceptibilities $k_{1}$, $k_{2}$, and $k_{3}$ (Eq. \ref{eq:K})
are equal to a constant $\chi$. In this case, the susceptibility
tensor $\mathbf{K}$ (Eq. \ref{eq:K}) assumes the particular form
\begin{equation}
\mathbf{K} = \chi \, \mathbf{I} \: ,
\label{eq:K-isotropic}
\end{equation}
where $\mathbf{I}$ represents the identity matrix.

@@

Let us pre-multiply the uniform internal field $\tilde{\mathbf{H}}^{\dagger}$
(Eq. \ref{eq:Hi-tilde}) by the transformed 
susceptibility tensor $\tilde{\mathbf{K}}$ 
(Eq. \ref{eq:H-tilde}) to obtain
\begin{equation}
\begin{split}
\tilde{\mathbf{M}} 
&= \tilde{\mathbf{K}} 
\left( \mathbf{I} - \tilde{\mathbf{N}}^{\dagger} \, \tilde{\mathbf{K}} \right)^{-1}
\tilde{\mathbf{H}}_{0} \\
&=  
\left( \mathbf{I} - \tilde{\mathbf{K}} \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\tilde{\mathbf{K}} \, \tilde{\mathbf{H}}_{0}
\end{split} \: ,
\label{eq:M-tilde}
\end{equation}
where $\tilde{\mathbf{M}}$
represents the transformed magnetization, as can be easily
verified by using Eqs. \ref{eq:M-KH} and \ref{eq:H-tilde}.
The matrix identity used for obtaining the second line of Eq. 
\ref{eq:M-tilde} is given by \citet[p. ~151]{searle1982}.
Equation \ref{eq:M-tilde} can be easily generalized for the case in 
which the ellipsoid has also a uniform remanent magnetization
$\tilde{\mathbf{M}}_{R}$. Let us first consider that
the uniform remanent magnetization satisfies the condition
$\tilde{\mathbf{H}}_{A} = \tilde{\mathbf{K}}^{-1} \tilde{\mathbf{M}}_{R}$,
where $\tilde{\mathbf{H}}_{A}$ represents a hypothetical 
uniform ancient field. Then, if we assume that $\tilde{\mathbf{H}}_{0}$,
in Eqs. \ref{eq:Hi-tilde} and \ref{eq:M-tilde},
is in fact the sum of the uniform magnetic field $\tilde{\mathbf{H}}_{0}$ 
and the hypothetical ancient field $\tilde{\mathbf{H}}_{A}$, we obtain the 
following generalized equation
\begin{equation}
\mathbf{M} = \mathbf{\Lambda}
\left(\mathbf{K} \, \mathbf{H}_{0} + \mathbf{M}_{R} \right) \: ,
\label{eq:M-remanence}
\end{equation}
where
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left( \mathbf{I} - \tilde{\mathbf{K}} \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\mathbf{V}^{\top} \: .
\label{eq:Lambda}
\end{equation}
Despite of the coordinate system transformation represented by the
matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
Eq. \ref{eq:M-remanence} is consistent with that given by
\citet[Eq. ~38]{clark1986}. It shows the combined effect of the
anisotropy of magnetic susceptibility (AMS) and the shape anisotropy.
The AMS is represented by the susceptibility tensor $\mathbf{K}$ 
(Eq. \ref{eq:K}) and reflects the
preferred orientation of the magnetic minerals forming the
body. The susceptibility tensor appears in Eq. \ref{eq:M-remanence},
defined in the main coordinate system (Fig. \ref{fig:orientation-angles}a),
and Eq. \ref{eq:Lambda}, defined in the local coordinate system
(Fig. \ref{fig:orientation-angles}b).
The shape anisotropy is represented, in Eq.\ref{eq:M-remanence} by the
depolarization tensor $\tilde{\mathbf{N}}^{\dagger}$ and
reflects the self-demagnetization associated to the body shape.
Notice that the resultant magnetization $\mathbf{M}$
(Eq. \ref{eq:M-remanence}) does not necessarily have
the same direction as the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}).
The angular difference $\theta$ between the resultant magnetization
and the inducing field depends on the combined effect of the
anisotropy of magnetic susceptibility and the shape anisotropy.

%The angular difference $\theta$ between the resultant magnetization
%and the inducing field is the same, whether the quantities
%are defined in the main coordinate system (Fig. \ref{fig:orientation-angles}a)
%or local coordinate system (Fig. \ref{fig:orientation-angles}b).
%For convenience, we define $\theta$ (in degrees) in the main coordinate system
%by using the scalar product definition as follows:
%\begin{equation}
%\theta = \frac{180}{\pi} \, \cos^{-1} \left( 
%\frac{\mathbf{M}^{\top}\mathbf{H}_{0}}
%{\| \mathbf{M} \| \; \| \mathbf{H}_{0} \|}\right) \: ,
%\label{eq:theta}
%\end{equation}
%where $\mathbf{M}$ is obtained by pre-multiplying the
%magnetization $\tilde{\mathbf{M}}$ (Eq. \ref{eq:M-tilde-remanence})
%by the orthogonal matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
%which is defined according to the ellipsoid type, and
%$\mathbf{H}_{0}$ is given by Eq. \ref{eq:H0}.
%
For the particular case in which the susceptibility is isotropic,
the susceptibility tensor is defined according to Eq. \ref{eq:K-isotropic}.
In this case, the magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-KH} and \ref{eq:M-remanence}),
referred to the main coordinate system (Fig. \ref{fig:orientation-angles}a),
and the matrix $\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda})
can be rewritten as follows:
\begin{equation}
\mathbf{M} = \mathbf{\Lambda} \left( \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \right) \: ,
\label{eq:M-K-isotropic}
\end{equation}
and
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left[ \begin{array}{ccc}
\frac{1}{1 - \chi \, \tilde{n}^{\dagger}_{11}} &  &  \\
 & \frac{1}{1 - \chi \, \tilde{n}^{\dagger}_{22}} &  \\
 &  & \frac{1}{1 - \chi \, \tilde{n}^{\dagger}_{33}}
\end{array} \right] \mathbf{V}^{\top} \: .
\label{eq:Lambda-K-isotropic}
\end{equation}
This equation is in perfect agreement with those presented 
by \citet[Eqs. ~13--15]{guo2001}.
The first term, depending on the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}), 
represents the induced magnetization whereas the term depending on
$\mathbf{M}_{R}$ is the remanent magnetization.
Equation \ref{eq:M-K-isotropic} reveals that, as pointed out by
many authors
\citep[e.g.,][]{maxwell1873, dubois1896, stoner1945, clark1986, sttraton2007},
the induced magnetization opposes the inducing field 
if it is parallel to an ellipsoid axis, 
independently of the ellipsoid type. Otherwise, the
magnetization is not necessarily parallel to
the inducing field.
If we additionally consider that $\chi << 1$, the matrix 
$\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda-K-isotropic})
approaches to the identity and the magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic})
can be approximated as follows:
\begin{equation}
\mathbf{M} \approx \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-approx}
\end{equation}
which is the classical equation describing the resultant
magnetization in applied geophysics \citep[p. ~89]{blakely1996}.
Notice that, in this particular case, the induced magnetization
is parallel to the inducing field $\mathbf{H}_{0}$ (Eq. \ref{eq:H0}),
whether it is parallel to an ellipsoid axis or not.
Usually, Eq. \ref{eq:M-approx}
is considered a good approximation for $\chi \leq 0.1$ SI.
